Efficient evaluation of threshold queries of derived fields in a numerical simulation database

### 2015-01-05 paper.tex
Fixed some typos.

@@

We impose a limit on the maximum number of locations that can be returned as a result of a threshold query in order to prevent having to return
the entire time-step or a significant fraction of it for queries with thresholds that are set too low. Currently this limit is around $10^6$ locations, which is
sufficient to examine a time-step in detail. In the case of the vorticity these locations contain all values within $\sim$8 times the root mean squared value of
the vorticity, which is about $25\%$ of the maximum. Given that we are interested in extreme events, obtaining the locations with values even within $50\%$ 
of the maximum would be sufficient. 
At the same time this also limits the amount of data that have to be returned to the user over the network as well as the amount of
data that have to be cached. Users receive an error message notifying them if their request has a threshold that is set too low. If a user
is interested in obtaining more data he or she can request the values of the derived field directly. Alternatively, if they are interested in the density distribution
of values they can examine the probability density function (e.g. Fig. \ref{fig:pdf}), which is computed using a similar strategy to threshold queries.

@@

In a system called the tree cache Lopez et al. \cite{Lopez} make use of a small application-aware cache to reduce access time to large datasets stored on
disk. The tree cache stores individual octants of octree datasets and exploits application-specific information to determine which octants to cache and to 
perform query reordering. This work has inspired the use of an application-aware cache for the evaluation of threshold queries. In contrast to the tree cache,
we do not cache raw data objects, but rather query results. Caching query results preserves the computational effort in addition to reducing I/O, which has a
much bigger impact on query performance and substantially reduces the size of the cache.. Additionally, the cache that we introduce resides on disk rather 
than in memory, which greatly increases its potential size. Lopez et al. also explore approximate querying through aggregation, which can be fairly easily
supported by our system but is of limited use as scientists performing threshold queries are usually interested in obtaining the exact locations where a field is 
at its highest values.

@@

The introduction of an application-aware cache for query results lays the ground work for the creation of a landmark database. Such a database can store the 
locations of the highest vorticity regions in the dataset or more broadly regions of interest and their associated statistics.

@@

\section{Acknowledgment}
The authors would like to thank the Turbulence Database Group at Johns Hopkins University as well as three anonymous reviewers for
their insightful comments and suggestions. This work is supported in part by the National Science Foundation
under Grants ....

>>>>

We impose a limit on the maximum number of locations that can be returned as a result of a threshold query in order to prevent having to return
the entire time-step or a significant fraction of it for queries with thresholds that are set too low. Currently this limit is around $10^6$ locations, which is
sufficient to examine a time-step in detail. In the case of the vorticity these locations contain all values above $\sim$8 times the root mean square value of
the vorticity, which is about $25\%$ of the maximum. Given that we are interested in extreme events, obtaining the locations with values even within $50\%$ 
of the maximum would be sufficient. 
At the same time this also limits the amount of data that have to be returned to the user over the network as well as the amount of
data that have to be cached. Users receive an error message notifying them if their request has a threshold that is set too low. If a user
is interested in obtaining more data he or she can request the values of the derived field directly. Alternatively, if they are interested in the density distribution
of values they can examine the probability density function (e.g. Fig. \ref{fig:pdf}), which is computed using a similar strategy to threshold queries.

@@

@@

In a system called the tree cache Lopez et al. \cite{Lopez} make use of a small application-aware cache to reduce access time to large datasets stored on
disk. The tree cache stores individual octants of octree datasets and exploits application-specific information to determine which octants to cache and to 
perform query reordering. This work has inspired the use of an application-aware cache for the evaluation of threshold queries. In contrast to the tree cache,
we do not cache raw data objects, but rather query results. Caching query results preserves the computational effort in addition to reducing I/O, which has a
much bigger impact on query performance and substantially reduces the size of the cache. Additionally, the cache that we introduce resides on disk rather 
than in memory, which greatly increases its potential size. Lopez et al. also explore approximate querying through aggregation, which can be fairly easily
supported by our system but is of limited use as scientists performing threshold queries are usually interested in obtaining the exact locations where a field is 
at its highest values.

@@

The introduction of an application-aware cache for query results lays the groundwork for the creation of a landmark database. Such a database can store the 
locations of the highest vorticity regions in the dataset or more broadly regions of interest and their associated statistics.

@@

\section{Acknowledgments}
The authors would like to thank the Turbulence Database Group at Johns Hopkins University as well as three anonymous reviewers for
their insightful comments and suggestions. This work is supported in part by the National Science Foundation
under Grants to..
